\section{Introduction}

In forested regions, the response of trees to solar irradiance, temperature, humidity, and subsurface moisture influences the timing of water flux to the atmosphere, of energy partitioning at the land surface, and of fixation of carbon [\textit{Bonan}, 2008]. In a Mediterranean climate, the season of high water supply is offset from the season of high atmospheric evaporative demand and high solar irradiance [\textit{Baldocchi and Xu}, 2007]. For forested landscapes in these climates, such as much of the Northern California Coast Range, the dominant tree species are evergreen [\textit{Woudenberg et al.}, 2010], yet their transpiration is not constant through the year [e.g. \textit{Vinukollu et al.}, 2011].  In this study, we demonstrate differences in transpiration seasonality between needleleaf and broadleaf evergreen trees; we show that these differences are due to different responses to surface soil moisture and atmospheric evaporative demand.

Water supply limitation can reduce evapotranspiration from the whole-plant scale (specific references are discussed below) up to the regional scale [\textit{Jung et al.}, 2010].  The root-zone moisture value at which different species become water-stressed can determine which species thrive in different hydrologic regimes [\textit{Rodriguez-Iturbe et al.}, 2001; \textit{Kumagai and Porporato}, 2012].  Some species, such as Douglas-fir [\textit{Granier}, 1987; \textit{Tan and Black}, 1976; \textit{Black}, 1979; \textit{Humphreys et al.}, 2003; \textit{Jassal et al.}, 2009], juniper [\textit{McDowell et al.}, 2008], lodgepole pine, limber pine, and subalpine fir [\textit{Pataki et al.}, 2000], Aleppo pine [\textit{Baquedano and Castillo}, 2006; \textit{Chirino et al.} 2011], and many species in the \textit{Pinaceae} family [\textit{Mart\'inez-Vilalta et al.}, 2004] reduce transpiration in response to relatively moderate soil water deficits.  This drought response strategy may protect the trees from hydraulic failure but reduce carbon uptake [\textit{McDowell et al.}, 2008].  In contrast, some species maintain high rates of transpiration even as the subsurface dries (e.g. pi\~non [\textit{McDowell et al.}, 2008], Kermes oak and Holm oak [\textit{Baquedano and Castillo}, 2006; \textit{Chirino et al.}, 2011; \textit{David et al.}, 2007], and a eucalyptus species (\textit{E. gomphocephala}) [\textit{Franks et al.}, 2007]); this drought response strategy may expose the trees to hydraulic failure if the drought is severe enough but allow them to continue fixing carbon [\textit{McDowell et al.}, 2008]. 

Transpiration also depends on the rate of stomatal closure in response to increasing atmospheric evaporative demand.  There is evidence of species differences in the response to increasing atmospheric evaporative demand.  Maximum stomatal conductance correlates with stomatal sensitivity to vapor pressure deficit (\textit{VPD}, kPa) [\textit{Oren et al.}, 1999].  This means that a species with high stomatal conductance at low and moderate \textit{VPD} (e.g. 1 kPa) also rapidly closes its stomata as \textit{VPD} increases, limiting the increase of transpiration at higher atmospheric evaporative demand.  In contrast, other species that have low maximum stomatal conductance, and thus lower transpiration at low \textit{VPD}, also have less stomatal closure with increasing \textit{VPD}.  In the Pacific northwestern U.S., two common conifers (\textit{Pseudotsuga menziesii} and \textit{Tsuga heterophylla} [\textit{Marshall and Waring}, 1984; \textit{Bond and Kavanagh}, 1999]) close their stomata more rapidly in response to increasing \textit{VPD} than do certain co-occurring broadleaf species (\textit{Acer circinatum} Pursh, \textit{Berberis nervosa} Pursh, \textit{Ceanothus velutinus} Dougl. ex Hook., \textit{Gaultheria shallon} Pursh, \textit{Rhododendron macrophyllum} G. Don, \textit{Castanopsis chrysophylla} (Dougl.) A.D.C., and \textit{Cornus nuttallii} Aud. ex T. \& G. [\textit{Marshall and Waring}, 1984]; \textit{Populus trichocarpa} Torr. \& Gray. and \textit{Alnus rubra} Bong. [\textit{Bond and Kavanagh}, 1999]).  Other cases of species differences in response to atmospheric evaporative demand have also been documented [\textit{Aranda et al.}, 2000; \textit{Mart\'inez-Vilalta et al.}, 2003].  Such differences in the relationship between stomatal conductance and atmospheric evaporative demand are not captured in common land surface models such as CLM [\textit{Oleson et al.}, 2010], which applies a simple linear relationship between relative humidity and stomatal conductance for all plant functional types.

In this paper, we investigate the seasonality of transpiration of five common evergreen tree species in a Mediterranean climate, and the dependence of transpiration on root zone water supply and atmospheric evaporative demand. We use intensive high-cadence (30-minute) observations of sap flow, meteorological conditions, and soil moisture to:
\begin{enumerate}
\item Demonstrate differences in seasonal patterns of transpiration among evergreen species located on the same hillslope.
\item Quantify the species differences in sensitivities to water supply and atmospheric demand that drive the differences in seasonal timing.
\item Show how these different sensitivities create species differences in synoptic-scale (daily to weekly) variability in transpiration.
\item Estimate the contributions of different species to regional transpiration in different seasons, and compare this estimate to remote-sensing-based (MODIS) estimates.
\end{enumerate}

Moreover, we apply statistical techniques that are not generally used in the sap flow literature (PCA/EOF, Markov chain Monte Carlo parameter estimation) to find patterns in a large, high-cadence, multi-year dataset.  We hope that these new ways of analyzing large volumes of sap flow observations can serve as a template for future analysis of large ecological datasets.